---
title: "Geomx_Data"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2025-01-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-libs, message = FALSE,  warning = FALSE, results = FALSE}
library(BulkSignalR)
library(igraph)
library(dplyr)
```


```{r}
output_dir <- 'C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/BulkSignalR'


```

```{r loading,eval=FALSE}

geomx_obj <- readRDS('geomx_qc_norm.RDS')

count_geomx = data.frame(geomx_obj@assayData$exprs) # count data

#q3norm_geomx = data.frame(geomx_obj@assayData$q3_norm) # q3 normalized data


```  


```{r seperating,eval=FALSE}
## seperating stroma and tumor AOIs

segment  = geomx_obj@phenoData@data$Segment

sample = geomx_obj@phenoData@data$Sample

combined = mapply(paste, sample, segment, MoreArgs = list(sep = "_"))

colnames(count_geomx) = combined 

col_ids_stroma <- which(grepl("stroma", colnames(count_geomx), ignore.case = TRUE))

# Create a new dataframe with only those columns
stroma_df <- count_geomx[, col_ids_stroma]

col_ids_tumor <- which(grepl("tumor", colnames(count_geomx), ignore.case = TRUE))

# Create a new dataframe with only those columns
tumor_df <- count_geomx[, col_ids_tumor]



```



step 01 : Prepare Dataset

```{r prepareDataset , eval=FALSE,cache=FALSE}
bsrdm_stroma <- prepareDataset(counts = stroma_df)


#bsrdm <- prepareDataset(counts = count_geomx, method = "TC")

# @param method: The normalization method ('UQ' (default) for upper quartile or 'TC' for total count. UQ.pc = 0.75. 

#bsrdm <- prepareDataset(counts = q3norm_geomx, normalize = FALSE, method = "q3_norm")

# print object
#str(bsrdm)

``` 


step 02 : learnParameters



```{r learnParameters , eval=FALSE,cache=FALSE,results=FALSE}
set.seed(123)

bsrdm <- learnParameters(bsrdm_stroma, 
         plot.folder = file.path(output_dir, 'figures/geomx/stroma_n_tumor/'), filename = "geomxUQ_stroma", verbose = TRUE)

# user can specify the null model from null.model = c("automatic", "mixedNormal", "normal", "kernelEmpirical","empirical", "stable")


# example

#bsrdm <- learnParameters(bsrdm, null.model = "normal",
#         plot.folder = "C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/BulkSignalR/figures/geomx", filename = "geomx_normal")
```




step 03 : Building a BSRInference object

```{r BSRInference , eval=FALSE,cache=FALSE}
bsrinf_stroma <- initialInference(bsrdm)


# parameters of initialInference method of bsrdm class

#setMethod("initialInference", "BSRDataModel", function(obj, rank.p=0.55,
#min.cor = 0.25, restrict.genes = NULL,
#reference=c("REACTOME-GOBP","REACTOME","GOBP"),
#max.pw.size=NULL, min.pw.size=NULL, min.positive=NULL, 
#use.full.network=FALSE, 
#restrict.pw=NULL,
#with.complex=NULL, 
#fdr.proc=c("BH","Bonferroni","Holm","Hochberg",                "SidakSS","SidakSD","BY","ABH","TSBH"))


LRinter.dataframe <- LRinter(bsrinf_stroma)

#head(LRinter.dataframe[order(LRinter.dataframe$qval <= 0.001),
  #c("L",  "R" ,"LR.corr", "pw.id",  "pw.name",  "pval"  , "qval")])
```




```{r BSRInferenceTfile, eval=FALSE}
write.table(LRinter.dataframe[order(LRinter.dataframe$qval <= 0.001),],
            file.path(output_dir, 'figures/geomx/stroma_n_tumor/geomx_LR_stroma.csv'),
            row.names = FALSE,
            sep = "\t",
            quote = FALSE)
```


```{r objectLoading, eval=FALSE}
# loading generated BulkSignalR objects

#bsrdm <- readRDS('C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/BulkSignalR/geomx_output/stroma_n_tumor/stroma/bsrdm_stroma.RDS')
bsrinf_stroma <- readRDS('C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/BulkSignalR/geomx_output/stroma_n_tumor/stroma/bsrinf_stroma.RDS')
```

## Reduction strategies

1. Reducing a BSRInference object to pathway

```{r ReduceToPathway, eval=FALSE,cache=FALSE}
     bsrinf.redP <- reduceToPathway(bsrinf)
```

2. Reducing a BSRInference object to best pathway


```{r ReduceToBestPathway, eval=FALSE,cache=FALSE}
     bsrinf.redBP    <- reduceToBestPathway(bsrinf_stroma)
```

3. Reducing to ligands or receptors

```{r ReduceToLigand, eval=FALSE,cache=FALSE}
     bsrinf.L    <- reduceToLigand(bsrinf)
     bsrinf.R    <- reduceToReceptor(bsrinf)
```

4. Combined reductions

```{r doubleReduction, eval=FALSE,cache=FALSE}
     bsrinf.redP  <- reduceToPathway(bsrinf)  
     bsrinf.redPBP <- reduceToBestPathway(bsrinf.redP) 
```

step 4 : Building a BSRSignature object


1. Scoring by ligand-receptor

```{r scoringLR  ,cache=FALSE,results=FALSE,eval=FALSE}

     bsrsig.redBP <- getLRGeneSignatures(bsrinf.redBP, qval.thres=0.001)

     scoresLR <- scoreLRGeneSignatures(bsrdm_stroma, bsrsig.redBP,
                                       name.by.pathway=FALSE)
     
     simpleHeatmap(scoresLR[1:10,], 
                   path = 'C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/BulkSignalR/figures/geomx/stroma_n_tumor/',
                   filename = "geomx_scoresLR_stroma",
                   column.names = TRUE, 
                   height = 5, width = 9,
                   pointsize = 10,
                   hcl.palette = "Cividis"
                   )
```

2. Scoring by pathway

```{r scoringPathway, cache=FALSE,results=FALSE,eval=FALSE}    
     bsrsig.redPBP <- getLRGeneSignatures(bsrinf.redPBP, qval.thres=0.01)
     
     scoresPathway <- scoreLRGeneSignatures(bsrdm, bsrsig.redPBP,
                                            name.by.pathway=TRUE)
     
     simpleHeatmap(scoresPathway[c(19,16,17,55,54,64),], 
                   path = 'C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/BulkSignalR/figures/geomx/stroma_n_tumor/',
                   filename = "geomx_scoresPathway_stroma",
                   column.names = TRUE, 
                   width = 9, 
                   height = 5, 
                   pointsize = 12,
                   hcl.palette = "Blue-Red 2"
                   )

```

## Other visualization utilities

1. Heatmap of ligand-receptor-target genes expression


```{r HeatmapRTL ,results=FALSE,,eval=FALSE}   
pathway1 <- "Interleukin-10 signaling"
signatureHeatmaps(
        pathway = pathway1,
        bsrdm = bsrdm,
        bsrsig = bsrsig.redPBP,
        path = 'C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/BulkSignalR/figures/geomx/stroma_n_tumor/',
        filename = "geomx_signatureheatmap_tumor",
        h.width  = 15,
        h.height = 10 ,
        show_column_names = TRUE)
```  

2. AlluvialPlot

```{r AlluvialPlot,results=FALSE,eval=FALSE}   

alluvialPlot(bsrinf_stroma,
              keywords = c("B2M"),
              type = "L",
              qval.thres = 0.001,
              path = 'C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/BulkSignalR/figures/geomx/stroma_n_tumor/',
              filename = "geomx_alluvial_stroma", 
              width  = 16, 
              height = 12
              )

```


2. BubblePlot


```{r BubblePlot,results=FALSE,eval=FALSE} 
pathway1 <- "Interleukin-21 signaling"
pathway2 <- "T cell receptor signaling pathway"
pathway3 <- "adaptive immune response"
pathway4 <- "Interferon gamma signaling"

pathways <- c(pathway1,pathway2,pathway3,pathway4)

bubblePlotPathwaysLR(bsrinf_stroma,
    pathways = pathways, 
    qval.thres  = 1,
    path = 'C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/BulkSignalR/figures/geomx/stroma_n_tumor/',
    color = "red",
    filename  = "geomx_bubble_stroma", 
    width  = 16, 
    height = 7,
    pointsize = 8
    #filter.L = c("ADAM12")
    #filter.R = c("ITGA3")
    ) 
```

3. Chordiagram

```{r Chordiagram,results=FALSE,eval=FALSE} 
chordDiagramLR (bsrinf_stroma,
                  path = 'C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/BulkSignalR/figures/geomx/stroma_n_tumor/',
                  filename = "geomx_chord_stroma",
                  pw.id.filter = "R-HSA-877300",
                  limit = 20,
                  width = 5, 
                  height = 4.5
    )

# R-HSA-6785807 : Interleukin-4 and Interleukin-13 signaling
# R-HSA-198933 : Immunoregulatory interactions between a Lymphoid
# R-HSA-877300 : Interferon gamma signaling
# GO:0050852 : T cell receptor signaling pathway
```


#### Network Analysis

```{r network1, eval=FALSE,cache=FALSE, results=FALSE} 
# Generate a ligand-receptor network and export it in .graphML
# for Cytoscape or similar tools
gLR <- getLRNetwork(bsrinf.redBP, qval.thres=1e-8)

# save to file
#write.graph(gLR,file="SDC-LR-network.graphml",format="graphml")

# As an alternative to Cytoscape, you can play with igraph package functions.
plot(gLR,
     edge.arrow.size=0.1,
     vertex.label.color="black",
     vertex.label.family="Helvetica",
     vertex.label.cex=0.1)

```



```{r network2, eval=FALSE,cache=FALSE, results=FALSE} 
# You can apply other functions.


# Community detection
u.gLR <- as.undirected(gLR) # most algorithms work for undirected graphs only
comm <- cluster_edge_betweenness(u.gLR)
# plot(comm,u.gLR,
#     vertex.label.color="black",
#     vertex.label.family="Helvetica",
#     vertex.label.cex=0.1)

# Cohesive blocks
cb <- cohesive_blocks(u.gLR)
plot(cb,u.gLR,
     vertex.label.color="black",
     vertex.label.family="Helvetica",
     vertex.label.cex=0.1,
     edge.color="black")
```



```{r network3, eval=FALSE,cache=FALSE, results=FALSE} 
# For the next steps, we just share the code below but graph generation function
# are commented to lighten the vignette.

# Generate a ligand-receptor network complemented with intra-cellular,
# receptor downstream pathways [computations are a bit longer here]
#
# You can save to a file for cystoscape or plot with igraph.

gLRintra <- getLRIntracellNetwork(bsrinf.redBP, qval.thres=1e-8, min.logFC=NULL))
# have to set min.logFC=NULL or else shoot an error

lay <- layout_with_kk(gLRintra)
# plot(gLRintra,
#     layout=lay,
#     edge.arrow.size=0.1,
#     vertex.label.color="black",
#     vertex.label.family="Helvetica",
#     vertex.label.cex=0.1)

# Reduce complexity by focusing on strongly targeted pathways
pairs <- LRinter(bsrinf.redBP)
top <- unique(pairs[pairs$pval<1e-20,c("pw.id","pw.name")])
#top
gLRintra.res <- getLRIntracellNetwork(bsrinf.redBP, qval.thres=0.01,restrict.pw=top$pw.id, min.logFC=NULL)
lay <- layout_with_fr(gLRintra.res)

# plot(gLRintra.res,
#     layout=lay,
#     edge.arrow.size=0.1,
#     vertex.label.color="black",
#     vertex.label.family="Helvetica",
#     vertex.label.cex=0.4)
    
```




### CellTypesToInteractions

```{r CellTypesToInteractions1, eval=FALSE}
 
#' # microenvironment cell populations
#' 
data(immune.signatures, package="BulkSignalR")
immune.signatures <- immune.signatures[immune.signatures$signature %in%
                                c("B cells","Dentritic cells","Macrophages",
                                  "NK cells","T cells","T regulatory cells"),]
data("tme.signatures", package="BulkSignalR")
signatures <- rbind(immune.signatures,tme.signatures[
          tme.signatures$signature%in%c("Endothelial cells","Fibroblasts"),])


```

```{r CellTypesToInteractions2, eval=FALSE}
tme.scores <- scoreSignatures(bsrdm, signatures)
#' 
#' # assignment
lr2ct <- assignCellTypesToInteractions(bsrdm, bsrinf, tme.scores)
```



Saving Objects: 




```{r saveobj, eval=FALSE}

saveRDS(bsrinf_stroma, file = "C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/BulkSignalR/geomx_output/stroma_n_tumor/bsrinf_stroma.RDS")
saveRDS(bsrdm, file = "C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/BulkSignalR/geomx_output/stroma_n_tumor/bsrdm_stroma.RDS")


```



```{r}
temp = bsrinf.redPBP@LRinter

patterns <- c("Interleukin", "Interferon", "B Cell", "T Cell")
pattern_regex <- paste(patterns, collapse = "|")

row_ids <- which(grepl(pattern_regex, temp$pw.name, ignore.case = TRUE))

filtered_pw <- temp[row_ids,]

df_IL_pw = data.frame(pw = filtered_pw$pw.name,qval = round(filtered_pw$qval,3))

knitr::kable(df_IL_pw)

```

```{r}
write.table(df_IL_pw,
            file.path(output_dir, 'figures/geomx/stroma_n_tumor/geomx_PW_tumor.csv'),
            row.names = FALSE,
            sep = "\t",
            quote = FALSE)
```



