---
title: "Geomx_normalized"
output: pdf_document
date: "2025-02-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("ggplot2")
library(gridExtra)
library(nichenetr)
library(Seurat)
#library(SeuratObject)
library(tidyverse)
```

# Nichenet for geomx data unnormalized

```{r loading data}

lr_network <- readRDS("C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/NichNet/lr_network_human_21122021.rds")

ligand_target_matrix <- readRDS("C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/NichNet/ligand_target_matrix_nsga2r_final.rds")

weighted_networks <- readRDS("C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/NichNet/weighted_networks_nsga2r_final.rds")

DEG_geomx = read.csv("C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/NichNet/DEG/dge_within_slide_Annotation_cell_bin_TRUE_CD8_Iba1_Segment_NACT_status.csv")

# normalized geomx data
norm_geomx = readRDS("C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/NichNet/deconvoluted_data/bp_res_mid_lvl_ct_45_expr_mtx_cleaned_norm.RDS")


lr_network <- lr_network %>% distinct(from, to)

```


```{r}
Macrophages_exp_norm = norm_geomx$Macrophages
Macrophages_exp_norm = t(Macrophages_exp_norm)
Tcell_exp_norm = norm_geomx$Tcells
Tcell_exp_norm = t(Tcell_exp_norm)
min(Tcell_exp_norm)
min(Macrophages_exp_norm)
```

######################################################

# exploring the expression data

```{r Tcells_plot}
df <- data.frame(expression = as.vector(unlist(Tcell_exp_norm)))

plot2 <- ggplot(df, aes(x = expression)) +
  geom_density(fill = "blue") +
  #xlim(1,2)+
  geom_vline(xintercept = 0, col = "red", linetype = "dashed") +
  geom_vline(xintercept = 1, col = "red", linetype = "dashed") +
  geom_vline(xintercept = 2, col = "red", linetype = "dashed") +
  theme_minimal()

plot2
```


```{r Macrophages_plot}
df <- data.frame(expression = as.vector(unlist(Macrophages_exp_norm)))

plot2 <- ggplot(df, aes(x = expression)) +
  geom_density(fill = "blue") +
  #xlim(1,2)+
  geom_vline(xintercept = 0, col = "red", linetype = "dashed") +
  geom_vline(xintercept = 1, col = "red", linetype = "dashed") +
  geom_vline(xintercept = 2, col = "red", linetype = "dashed") +
  theme_minimal()

plot2
```

######################

# Filtering the expression data

```{r}
x = apply(Macrophages_exp_norm, 2, function(x) mean(x <= 1.491))

hist(x)
```


```{r}
x = apply(Tcell_exp_norm, 2, function(x) mean(x <= 1.788))

hist(x)
```

According to the vignettes we can select the genes that have 10% of non zero values across samples as in single cell data. which means in our cases the threshold would be 0.90

setting the threshold to 0 implies that none of the remaining genes after filtering contains zero values in any of the samples . And this would remove the whole first peak.

```{r}
Macrophages_ID <- apply(Macrophages_exp_norm, 2, function(x) mean(x <= 1.491) <= 0)

#15094 - sum(Macrophages_ID)

sum(Macrophages_ID)

Tcells_ID <- apply(Tcell_exp_norm, 2, function(x) mean(x <= 1.788) <= 0)

#15094 - sum(Tcells_ID)

sum(Tcells_ID)

```

```{r}
Macrophages_exp_norm_filtered <- Macrophages_exp_norm[,Macrophages_ID]
Tcell_exp_norm_filtered <- Tcell_exp_norm[,Tcells_ID]
```

```{r}

df <- data.frame(expression = as.vector(unlist(Macrophages_exp_norm)))

plot3 <- ggplot(df, aes(x = expression)) +
  geom_density(fill = "blue") +
  geom_vline(xintercept = 0, col = "red", linetype = "dashed") +
  geom_vline(xintercept = 1, col = "red", linetype = "dashed") +
  geom_vline(xintercept = 2, col = "red", linetype = "dashed") +
  ggtitle("Macrophage(normalized counts)") +
  theme_minimal()

plot3
```
```{r}
df <- data.frame(expression = as.vector(unlist(Macrophages_exp_norm_filtered)))

plot3 <- ggplot(df, aes(x = expression)) +
  geom_density(fill = "blue") +
  geom_vline(xintercept = 0, col = "red", linetype = "dashed") +
  geom_vline(xintercept = 1, col = "red", linetype = "dashed") +
  geom_vline(xintercept = 2, col = "red", linetype = "dashed") +
  ggtitle("Macrophage(normalized counts filtred: threshold = 0.90)") +
  annotate("text", x = 9, y = 0.25, label = "No of genes = 11545", size = 3, color = "black") +
  theme_minimal()

plot3
```




```{r}

df <- data.frame(expression = as.vector(unlist(Tcell_exp_norm_filtered)))

plot3 <- ggplot(df, aes(x = expression)) +
  geom_density(fill = "blue") +
  ggtitle("Tcells(normalized counts filtred: threshold = 0.00)") +
  geom_vline(xintercept = 0, col = "red", linetype = "dashed") +
  geom_vline(xintercept = 1, col = "red", linetype = "dashed") +
  geom_vline(xintercept = 2, col = "red", linetype = "dashed") +
  theme_minimal()

plot3
```


```{r}
df <- data.frame(expression = as.vector(unlist(Macrophages_exp_norm_filtered)))

plot3 <- ggplot(df, aes(x = expression)) +
  geom_density(fill = "blue") +
  ggtitle("Macrophage(normalized counts filtred: threshold = 0)") +
  geom_vline(xintercept = 0, col = "red", linetype = "dashed") +
  geom_vline(xintercept = 1, col = "red", linetype = "dashed") +
  geom_vline(xintercept = 2, col = "red", linetype = "dashed") +
  theme_minimal()

plot3
```


```{r}
Macrophages_exp_norm_log2 = log2(Macrophages_exp_norm)

```


```{r}
Macrophages_ID <- apply(Macrophages_exp_norm_log2 , 2, function(x) mean(x) <= 1)

#15094 - sum(Macrophages_ID)

sum(Macrophages_ID)

#Tcells_ID <- apply(Tcell_exp_norm, 2, function(x) mean(x <= 1.788) <= 0)

#15094 - sum(Tcells_ID)

#sum(Tcells_ID)

```

```{r}
Macrophages_exp_norm_log2_filtered <- Macrophages_exp_norm_log2[,!Macrophages_ID]

df <- data.frame(expression = as.vector(Macrophages_exp_norm_log2_filtered))

plot3 <- ggplot(df, aes(x = expression)) +
  geom_density(fill = "blue") +
  ggtitle("Macrophage(normalized counts filtred: threshold = 0)") +
  geom_vline(xintercept = 0, col = "red", linetype = "dashed") +
  #geom_vline(xintercept = 1, col = "red", linetype = "dashed") +
  #geom_vline(xintercept = 2, col = "red", linetype = "dashed") +
  theme_minimal()

plot3
```


```{r}
summary(Tcell_exp_norm[,1:5])
```

```{r}
summary(Tcell_exp_norm_filtered[,1:5])
```




########################

# LR Analysis for both with and Without filtering

1. without filtering use "Macrophages_exp_norm" and "Tcell_exp_norm"
2. with filtering use "Macrophages_exp_norm_filtered" and "Tcell_exp_norm_filtered"

```{r}
colnames(Macrophages_exp_norm_filtered) <- convert_alias_to_symbols(colnames(Macrophages_exp_norm_filtered), "human", verbose = FALSE)


colnames(Tcell_exp_norm_filtered) <- convert_alias_to_symbols(colnames(Tcell_exp_norm_filtered), "human", verbose = FALSE)

```

sender = Macrophages
Receiver = Tcells

```{r}
ligands <- lr_network %>% pull(from) %>% unique()
expressed_ligands <- intersect(ligands,colnames(Macrophages_exp_norm_filtered))

receptors <- lr_network %>% pull(to) %>% unique()
expressed_receptors <- intersect(receptors,colnames(Tcell_exp_norm_filtered))

potential_ligands <-  lr_network %>% filter(from %in% expressed_ligands & to %in% expressed_receptors) %>% 
  pull(from) %>% unique()

```

background_expressed_genes comes from the receiver

```{r}
background_expressed_genes <- colnames(Tcell_exp_norm_filtered) %>% .[. %in% rownames(ligand_target_matrix)]
```


# 1. Gene set of interests: Differentially expressed genes


```{r}
gene = DEG_geomx[DEG_geomx$Pr...t.. < 0.05,]$Gene


geneset_oi <- gene %>% .[. %in% rownames(ligand_target_matrix)] 

```

```{r}
intersect(geneset_oi,background_expressed_genes)
intersect(geneset_oi,potential_ligands)
```

# 2. Highly Variable genes


```{r}

seuratObj_from_geomx2 = readRDS("C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/NichNet/seuratObj_from_norm_geomx2.RDS")

#HVGs = VariableFeatures(seuratObj_from_geomx2)
HVGs_top100 = head(VariableFeatures(seuratObj_from_geomx2), 100)

geneset_oi <- HVGs_top100 %>% .[. %in% rownames(ligand_target_matrix)] 
```

# 3. highly_expressed_genes

```{r}
avg_expression <- colMeans(tt)

highly_expressed_genes_tcells <- names(sort(avg_expression, decreasing = TRUE)[1:100])

geneset_oi <- highly_expressed_genes_tcells %>% .[. %in% rownames(ligand_target_matrix)]
```


# predicting ligand activity

```{r}
ligand_activities <- predict_ligand_activities(geneset = geneset_oi,
                                               background_expressed_genes = background_expressed_genes,
                                               ligand_target_matrix = ligand_target_matrix,
                                               potential_ligands = potential_ligands)
```

```{r}
ligand_activities <- ligand_activities %>% arrange(-aupr_corrected) %>%
  mutate(rank = rank(desc(aupr_corrected)))
```

```{r}
saveRDS(ligand_activities, file = "C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/NichNet/LR_activities_norm_filtered_threshold_0.RDS")

```


```{r}
best_upstream_ligands <- ligand_activities %>% top_n(30, aupr_corrected) %>%
  arrange(-aupr_corrected) %>% pull(test_ligand)
```

```{r}
active_ligand_target_links_df <- best_upstream_ligands %>%
  lapply(get_weighted_ligand_target_links,
         geneset = geneset_oi,
         ligand_target_matrix = ligand_target_matrix,
         n = 200) %>% bind_rows() # top 200 targets


active_ligand_target_links <- prepare_ligand_target_visualization(
  ligand_target_df = active_ligand_target_links_df,
  ligand_target_matrix = ligand_target_matrix,
  cutoff = 0.25) # weight quantile cutoff

order_ligands <- intersect(best_upstream_ligands, colnames(active_ligand_target_links)) %>% rev() # rev reverse the order

order_targets <- active_ligand_target_links_df$target %>% unique() %>% intersect(rownames(active_ligand_target_links))

vis_ligand_target <- t(active_ligand_target_links[order_targets,order_ligands])


```

```{r}
saveRDS(active_ligand_target_links_df, file = "C:/Users/Sahas/Downloads/Masters_Thesis/Ligand-receptor/NichNet/Lactive_ligand_target_links_df_norm_threshold_0.RDS")

```

```{r, fig.width=15, fig.height=6}
p_ligand_target_network <- make_heatmap_ggplot(vis_ligand_target, "Prioritized Macrophage-ligands", "genes in Tcells",
                    color = "purple", legend_title = "Regulatory potential") +
  scale_fill_gradient2(low = "whitesmoke",  high = "purple")+
  ggtitle("Macrophage-T Cell (Gene set of Interest = HVGs):Norm filtered ( <= 0)")


p_ligand_target_network <- p_ligand_target_network +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))

p_ligand_target_network
```

# Which receptors of the receiver cell population (T cells) can potentially bind to the prioritized ligands from the sender cell population (Macrophages).

```{r}

ligand_receptor_links_df <- get_weighted_ligand_receptor_links(
  best_upstream_ligands, expressed_receptors,
  lr_network, weighted_networks$lr_sig) 

vis_ligand_receptor_network <- prepare_ligand_receptor_visualization(
  ligand_receptor_links_df,
  best_upstream_ligands,
  order_hclust = "both") 

(make_heatmap_ggplot(t(vis_ligand_receptor_network), 
                     y_name = "Prioritized Macrophage-ligands", x_name = "Receptors expressed by T cells",  
                     color = "mediumvioletred", legend_title = "Prior interaction potential"))
```


# Top 5 ligands normalized expressions

```{r}
ligand_IDS = intersect(best_upstream_ligands, colnames(active_ligand_target_links))
ligand_expr_norm_0 = Macrophages_exp_norm[,ligand_IDS]
summary(ligand_expr_norm_0[,1:5])
```

```{r, fig.width=15, fig.height=6}
df1 <- data.frame(expression = as.vector(unlist(ligand_expr_norm[,1])))
df2 <- data.frame(expression = as.vector(unlist(ligand_expr_norm[,2])))
df3 <- data.frame(expression = as.vector(unlist(ligand_expr_norm[,3])))
df4 <- data.frame(expression = as.vector(unlist(ligand_expr_norm[,4])))
df5 <- data.frame(expression = as.vector(unlist(ligand_expr_norm[,5])))

plot1 <- ggplot(df1, aes(x = expression)) +
  geom_density(fill = "blue") +
  #geom_vline(xintercept = 1, col = "red", linetype = "dashed") +
  #geom_vline(xintercept = 2, col = "red", linetype = "dashed") +
  ggtitle("Potential Ligand: TGM2 (normalized counts)") +
  theme_minimal()



plot2 <- ggplot(df2, aes(x = expression)) +
  geom_density(fill = "blue") +
  #geom_vline(xintercept = 1, col = "red", linetype = "dashed") +
  #geom_vline(xintercept = 2, col = "red", linetype = "dashed") +
  ggtitle("Potential Ligand:  APP(normalized counts)") +
  theme_minimal()


plot3 <- ggplot(df3, aes(x = expression)) +
  geom_density(fill = "blue") +
  #geom_vline(xintercept = 1, col = "red", linetype = "dashed") +
  #geom_vline(xintercept = 2, col = "red", linetype = "dashed") +
  ggtitle("Potential Ligand:  HLA-A(normalized counts)") +
  theme_minimal()



plot4 <- ggplot(df4, aes(x = expression)) +
  geom_density(fill = "blue") +
  #geom_vline(xintercept = 1, col = "red", linetype = "dashed") +
  #geom_vline(xintercept = 2, col = "red", linetype = "dashed") +
  ggtitle("Potential Ligand:   ADAM10 (normalized counts)") +
  theme_minimal()

plot5 <- ggplot(df5, aes(x = expression)) +
  geom_density(fill = "blue") +
  #geom_vline(xintercept = 1, col = "red", linetype = "dashed") +
  #geom_vline(xintercept = 2, col = "red", linetype = "dashed") +
  ggtitle("Potential Ligand:  HLA-DQB1 (normalized counts)") +
  theme_minimal()

#grid.arrange(plot1, plot2, plot3, plot4, ncol = 2, nrow =2)
grid.arrange(plot1, plot3, nrow = 2)
grid.arrange(plot2, plot4, plot5, nrow = 3)
```



